
---
title: "R Notebook" 
output: html_document
---

**Sports analytics** exercise based on *Moneyball*—the story of how the Oakland A’s used data to build a winning baseball team on a tight budget. Let's break this down in simple terms and explain the **purpose** of each code section from Videos 2 and 3.


## 🎯 **Overall Purpose of the Exercise**

This project is teaching you how to:

1. **Use historical baseball data** to understand what actually contributes to winning games.
2. **Build a data-driven strategy** to pick players who are undervalued but effective.
3. **Apply linear regression** to figure out what statistics (like OBP or SLG) truly matter in predicting success (runs scored, runs allowed, and wins).

This mirrors what the Oakland A’s did in real life: **use statistics to find hidden value** in players that big-budget teams ignored.

---

###  **Load and Explore the Dataset**
# VIDEO 2

https://ocw.mit.edu/courses/15-071-the-analytics-edge-spring-2017/pages/linear-regression/moneyball-the-power-of-sports-analytics/video-2-making-it-to-the-playoffs/

```{r}
baseball = read.csv("baseball.csv")
str(baseball)
```
You're loading a dataset of MLB team stats from different years. This includes:

* **Runs Scored (RS)** and **Runs Allowed (RA)**
* **On Base Percentage (OBP)** and **Slugging Percentage (SLG)**
* **Wins (W)** and whether they made the **Playoffs**

*Goal*: Understand what data we have.

---

### 📌 **Focus on Moneyball Years**

```{r}
# Create a new dataset containing only seasons before 2002 (Moneyball era)
moneyball = subset(baseball, Year < 2002)

# Check the structure of the Moneyball subset
str(moneyball)
```

 You’re narrowing it down to the **Moneyball era** (before 2002), because that’s when Billy Beane used analytics to change how players were selected.

*Goal*: Analyze only the data that was available at the time the A’s were making decisions.

---
###  **Create Run Differential (RD)**

```{r}
# Calculate Run Differential (RD = Runs Scored - Runs Allowed)
# Teams that score more than they allow tend to win more games
moneyball$RD = moneyball$RS - moneyball$RA

# Confirm the new variable was added correctly
str(moneyball)
```

Narrowing it down to the **Moneyball era** (before 2002), because that’s when Billy Beane used analytics to change how players were selected.

*Goal*: Analyze only the data that was available at the time the A’s were making decisions.

###  **Create Run Differential (RD)**

```{r}
# Calculate Run Differential (RD = Runs Scored - Runs Allowed)
# Teams that score more than they allow tend to win more games
moneyball$RD = moneyball$RS - moneyball$RA

# Confirm the new variable was added correctly
str(moneyball)

```

RD = **Runs Scored - Runs Allowed**
If a team scores more than it allows, it's more likely to win.

*Goal*: Create a new variable that might help explain team success (Wins).

---

###  **Check Relationship: Run Differential vs Wins**

```{r}
# Create a scatterplot to see if there's a relationship between RD and Wins
# A straight-line trend would suggest we can use linear regression
plot(moneyball$RD, moneyball$W)

```
creating a scatterplot to see if teams that score more than they allow also win more games.

 *Goal*: Visualize whether there’s a **linear relationship** (straight-line pattern) between RD and Wins.

### **Build First Model: Predict Wins from RD**

```{r}
# Build a linear regression model to predict Wins using Run Differential
WinsReg = lm(W ~ RD, data=moneyball)

# Show the results of the regression
summary(WinsReg)

```
We are building a **linear regression model**:

* Inputs: Run Differential (RD)
* Output: Wins (W)

Model says:

> For every **1 extra run of RD**, a team wins **0.106 more games** on average.

📈 R-squared = 0.88 → Very strong model (88% of wins can be explained by RD).

 *Goal*: Prove that RD is a powerful predictor of wins.


