---
title: "R Notebook"
output: html_notebook
---


```{r}
2-3

```

```{r}
# Division
2/3
```

```{r}
# Exponentiation
2^3 
```
```{r}
# Square root
sqrt(2)
```

```{r}
# Logarithms
log(2)
```
```{r}
#Question_1: Compute the log base 5 of 10 and the log of 10.
# Log base 5 of 10
log_base5_of_10 <- log(10, base = 5)
log_base5_of_10
```

```{r}
# Log base 10 of 10 (common log)
log_base10_of_10 <- log10(10)
log_base10_of_10
```
###Computing some offensive metrics in Baseball

```{r}
#Batting Average=(No. of Hits)/(No. of At Bats)
#What is the batting average of a player that bats 29 hits in 112 at bats?
BA=(29)/(112)
BA
```

```{r}
Batting_Average=round(BA,digits = 3)
Batting_Average
```

```{r}
#Question_2:What is the batting average of a player that bats 42 hits in 212 at bats?

hits <- 42
at_bats <- 212

batting_average <- hits / at_bats

# Print result rounded to 3 decimal places
cat("Batting Average =", round(batting_average, 3), "\n")

```
```{r}
# Define the stats
AB <- 515   # At Bats
H  <- 172   # Hits
BB <- 84    # Walks (Bases on Balls)
HBP <- 5    # Hit by Pitch
SF <- 6     # Sacrifice Flies

# Calculate On Base Percentage (OBP)
OBP <- (H + BB + HBP) / (AB + BB + HBP + SF)

# Print result rounded to 3 decimals
cat("On Base Percentage (OBP) =", round(OBP, 3), "\n")

```
# Question_3:Compute the OBP for a player with the following general stats:
AB=565, H=156, BB=65, HBP=3, SF=7


```{r}

# Define the stats
AB<-565
H<-156
BB<-65
HBP<-3
SF<-7

# Calculate On Base Percentage (OBP)
OBP <- (H + BB + HBP) / (AB + BB + HBP + SF)

# Print result rounded to 3 decimals
cat("On Base Percentage (OBP) =", round(OBP, 3), "\n")


```
### Often you will want to test whether something is less than, greater than or equal to something.

```{r}
3 == 8# Does 3 equals 8?
```
```{r}
3 != 8# Is 3 different from 8?
```

```{r}
3>4
```

###The logical operators are & for logical AND, | for logical OR, and ! for NOT. These are some examples:
```{r}
# Logical Disjunction (or)

FALSE | FALSE # False OR False
```
```{r}
# Logical Conjunction (and)
TRUE & FALSE #True AND False
```

```{r}
# Negation
! FALSE # Not False
```
```{r}
# Combination of statements
2 < 3 | 1 == 5 # 2<3 is True, 1==5 is False, True OR False is True
```

# Assigning Values to Variables

In R, you create a variable and assign it a value using <- as follows
```{r}
Total_Bases <- 6 + 5
Total_Bases*3
```
To see the variables that are currently defined, use ls (as in “list”)

```{r}
ls()
```