---
title: "R Notebook Activity "
output: html_notebook
---


```{r}
2-3

```

```{r}
# Division
2/3
```

```{r}
# Exponentiation
2^3 
```
```{r}
# Square root
sqrt(2)
```

```{r}
# Logarithms
log(2)
```
```{r}
#Question_1: Compute the log base 5 of 10 and the log of 10.
# Log base 5 of 10
log_base5_of_10 <- log(10, base = 5)
log_base5_of_10
```

```{r}
# Log base 10 of 10 (common log)
log_base10_of_10 <- log10(10)
log_base10_of_10
```
###Computing some offensive metrics in Baseball

```{r}
#Batting Average=(No. of Hits)/(No. of At Bats)
#What is the batting average of a player that bats 29 hits in 112 at bats?
BA=(29)/(112)
BA
```

```{r}
Batting_Average=round(BA,digits = 3)
Batting_Average
```

```{r}
#Question_2:What is the batting average of a player that bats 42 hits in 212 at bats?

hits <- 42
at_bats <- 212

batting_average <- hits / at_bats

# Print result rounded to 3 decimal places
cat("Batting Average =", round(batting_average, 3), "\n")

```
```{r}
# Define the stats
AB <- 515   # At Bats
H  <- 172   # Hits
BB <- 84    # Walks (Bases on Balls)
HBP <- 5    # Hit by Pitch
SF <- 6     # Sacrifice Flies

# Calculate On Base Percentage (OBP)
OBP <- (H + BB + HBP) / (AB + BB + HBP + SF)

# Print result rounded to 3 decimals
cat("On Base Percentage (OBP) =", round(OBP, 3), "\n")

```
# Question_3:Compute the OBP for a player with the following general stats:
AB=565, H=156, BB=65, HBP=3, SF=7


```{r}

# Define the stats
AB<-565
H<-156
BB<-65
HBP<-3
SF<-7

# Calculate On Base Percentage (OBP)
OBP <- (H + BB + HBP) / (AB + BB + HBP + SF)

# Print result rounded to 3 decimals
cat("On Base Percentage (OBP) =", round(OBP, 3), "\n")


```
### Often you will want to test whether something is less than, greater than or equal to something.

```{r}
3 == 8# Does 3 equals 8?
```
```{r}
3 != 8# Is 3 different from 8?
```

```{r}
3>4
```

###The logical operators are & for logical AND, | for logical OR, and ! for NOT. These are some examples:
```{r}
# Logical Disjunction (or)

FALSE | FALSE # False OR False
```
```{r}
# Logical Conjunction (and)
TRUE & FALSE #True AND False
```

```{r}
# Negation
! FALSE # Not False
```
```{r}
# Combination of statements
2 < 3 | 1 == 5 # 2<3 is True, 1==5 is False, True OR False is True
```

# Assigning Values to Variables

In R, you create a variable and assign it a value using <- as follows
```{r}
Total_Bases <- 6 + 5
Total_Bases*3
```
To see the variables that are currently defined, use ls (as in “list”)

```{r}
ls()
```
To delete a variable, use rm (as in “remove”)
```{r}
rm(Total_Bases)
```

Either <- or = can be used to assign a value to a variable, but I prefer <- because is less likely to be confused with the logical operator ==

# Vectors

The basic type of object in R is a vector, which is an ordered list of values of the same type. You can create a vector using the c() function (as in “concatenate”).

```{r}
pitches_by_innings <- c(12, 15, 10, 20, 10) 
pitches_by_innings
```

```{r}
strikes_by_innings <- c(9, 12, 6, 14, 9)
strikes_by_innings
```
# Question_4: Define two vectors,runs_per_9innings and hits_per_9innings, each with five elements. 

```{r}
# Runs allowed per 9 innings for five pitchers/teams
runs_per_9innings <- c(3, 4, 2, 3, 4)
runs_per_9innings

# Hits allowed per 9 innings for the same five pitchers/teams
hits_per_9innings <- c(7, 8, 6, 7, 8)
hits_per_9innings
```
There are also some functions that will create vectors with regular patterns, like repeated elements.

```{r}
# replicate function
rep(2, 5)
```
```{r}
rep(1,4)
```
```{r}
# consecutive numbers
1:5
```
```{r}
2:10
```
```{r}
# sequence from 1 to 10 with a step of 2
seq(1, 10, by=2)
```
```{r}
seq(2,13,by=3)
```
Many functions and operators like + or - will work on all elements of the vector.

```{r}
# add vectors
pitches_by_innings+strikes_by_innings
```
```{r}
# compare vectors
pitches_by_innings == strikes_by_innings
```
```{r}
# find length of vector
length(pitches_by_innings)
```
```{r}
# find minimum value in vector
min(pitches_by_innings)
```

```{r}
# find average value in vector
mean(pitches_by_innings)
```
You can access parts of a vector by using [. Recall what the value is of the vector pitches_by_innings.

```{r}
```


```{r}
pitches_by_innings
```

```{r}
# If you want to get the first element:
pitches_by_innings[1]
```
#Question_5: Get the first element of hits_per_9innings.

```{r}
hits_per_9innings[1]

```

If you want to get the last element of pitches_by_innings without explicitly typing the number of elements of pitches_by_innings, make use of the length function, which calculates the length of a vector:
```{r}
pitches_by_innings[length(pitches_by_innings)]
```

# Question_6: Get the last element of hits_per_9innings.
```{r}
length(hits_per_9innings)

```
```{r}
hits_per_9innings[5]
```
***

You can also extract multiple values from a vector. For instance to get the 2nd through 4th values use

```{r}
pitches_by_innings[c(2, 3, 4)]
```
Vectors can also be strings or logical values

```{r}
player_positions <- c("catcher", "pitcher", "infielders", "outfielders")
player_positions
```

# Data Frames

In statistical applications, data is often stored as a data frame, which is like a spreadsheet, with rows as observations and columns as variables. To manually create a data frame, use the data.frame() function

```{r}
data.frame(bonus = c(2, 3, 1),#in millions 
           active_roster = c("yes", "no", "yes"), 
           salary = c(1.5, 2.5, 1))#in millions 
```

Most often you will be using data frames loaded from a file. For example, load the results of a fan’s survey. The function load or read.table can be used for this.

# How to Make a Random Sample
To randomly select a sample use the function sample(). The following code selects 5 numbers between 1 and 10 at random (without duplication)


```{r}
sample(1:10, size=5)
```
```{r}
x<-c("yes", "no", "no", "yes", "yes", "no", "yes", "yes","yes")
x
table(x)
```



The first argument gives the vector of data to select elements from.
The second argument (size=) gives the size of the sample to select.
Taking a simple random sample from a data frame is only slightly more complicated, having two steps:

Use sample() to select a sample of size n from a vector of the row numbers of the data frame.
Use the index operator [ to select those rows from the data frame.
Consider the following example with fake data. First, make up a data frame with two columns. (LETTERS is a character vector of length 26 with capital letters âAâ to âZâ; LETTERS is automatically defined and pre-loaded in R)

```{r}
bar <- data.frame(var1 = LETTERS[1:10], var2 = 1:10)
# Check data frame
bar
```

Suppose you want to select a random sample of size 5. First, define a variable n with the size of the sample, i.e. 5
```{r}
n <- 5
```

Now, select a sample of size 5 from the vector with 1 to 10 (the number of rows in bar). Use the function nrow() to find the number of rows in bar instead of manually entering that number.

Use : to create a vector with all the integers between 1 and the number of rows in bar.

```{r}
samplerows <- sample(1:nrow(bar), size=n) 
# print sample rows
samplerows
```
The variable samplerows contains the rows of bar which make a random sample from all the rows in bar. Extract those rows from bar with

```{r}
# extract rows
barsample <- bar[samplerows, ]
# print sample
print(barsample)
```
The code above creates a new data frame called barsample with a random sample of rows from bar.

In a single line of code:

```{r}
bar[sample(1:nrow(bar), n), ]
```
# Using Tables
The table() command allows us to look at tables. Its simplest usage looks like table(x) where x is a categorical variable.

For example, a survey asks people if they support the home team or not. The data is

Yes, No, No, Yes, Yes

We can enter this into R with the c() command, and summarize with the table() command as follows

```{r}
x <- c("Yes","No","No","Yes","Yes")
table(x)
```
# Numerical measures of center and spread
Suppose, MLB Teams’ CEOs yearly compensations are sampled and the following are found (in millions)

12 .4 5 2 50 8 3 1 4 0.25

```{r}
sals <- c(12, .4, 5, 2, 50, 8, 3, 1, 4, 0.25)
# the average
mean(sals) 
```
```{r}
# the variance
var(sals)
```
```{r}
# the standard deviation
sd(sals)
```
```{r}
# the median
median(sals)
```

```{r}
# Tukey's five number summary, usefull for boxplots
# five numbers: min, lower hinge, median, upper hinge, max
fivenum(sals)
```

```{r}
# summary statistics
summary(sals)
```
# How about the mode?
In R we can write our own functions, and a first example of a function is shown below in order to compute the mode of a vector of observations x

```{r}
# Function to find the mode, i.e. most frequent value
getMode <- function(x) {
     ux <- unique(x)
     ux[which.max(tabulate(match(x, ux)))]
 }
```


As an example, we can use the function defined above to find the most frequent value of the number of pitches_by_innings

```{r}
# Most frequent value in pitches_by_innings
getMode(pitches_by_innings)
```

# Question_7: Find the most frequent value of hits_per_9innings.

```{r}
#Question_7: Find the most frequent value of hits_per_9innings.

getMode(hits_per_9innings)
```

#Question_8: Summarize the following survey with the `table()` command:
#What is your favorite day of the week to watch baseball? A total of 10 fans submitted this survey.
#Saturday, Saturday, Sunday, Monday, Saturday,Tuesday, Sunday, Friday, Friday, Monday

```{r}
Favorite_day_of_week <- c("Saturday", "Saturday", "Sunday", "Monday", "Saturday","Tuesday","Sunday","Friday","Friday","Monday")
table(Favorite_day_of_week)

```

# Question_9: What is the most frequent answer recorded in the survey? Use the getMode function to compute results.

```{r}
getMode(Favorite_day_of_week)
```


