---
title: 'Module 3 : Activities 7 and 8'
output:
  html_notebook: default
  pdf_document: default
  html_document:
    df_print: paged
---



```{r}
# VIDEO 2


# Read in data
baseball = read.csv("baseball.csv")
str(baseball)
```
```{r}
# Create a new dataset containing only seasons before 2002 (Moneyball era)
moneyball = subset(baseball, Year < 2002)

# Check the structure of the Moneyball subset
str(moneyball)
```

```{r}
# Calculate Run Differential (RD = Runs Scored - Runs Allowed)
# Teams that score more than they allow tend to win more games
moneyball$RD = moneyball$RS - moneyball$RA

# Confirm the new variable was added correctly
str(moneyball)
```
RD = **Runs Scored - Runs Allowed**
If a team scores more than it allows, it's more likely to win.

```{r}
# Create a scatterplot to see if there's a relationship between RD and Wins
# A straight-line trend would suggest we can use linear regression
plot(moneyball$RD, moneyball$W)

```

```{r}
# Build a linear regression model to predict Wins using Run Differential
WinsReg = lm(W ~ RD, data=moneyball)

# Show the results of the regression
summary(WinsReg)
```
We are building a **linear regression model**:

* Inputs: Run Differential (RD)
* Output: Wins (W)

Model says:

> For every **1 extra run of RD**, a team wins **0.106 more games** on average.

📈 R-squared = 0.88 → Very strong model (88% of wins can be explained by RD).

 *Goal*: Prove that RD is a powerful predictor of wins.

###If a baseball team scores 763 runs and allows 614 runs, how many games do we expect the team to win?

Using the linear regression model constructed during the lecture, enter the number of games we expect the team to win:

```{r}
rs = 763
ra = 614

rd = rs - ra
rd
```

```{r}
wins =80.881 + 0.105766 * rd
wins
```

## According to our model if the A's score 763 runs and allow 614 runs,  would win around 96 or 97 games

---

```{r}
# Video 3
# Confirm the structure of the data again (double-check)
str(moneyball)
```

```{r}
# Build a regression model to predict Runs Scored (RS)
# Based on three batting stats: OBP, SLG, and BA
RunsReg = lm(RS ~ OBP + SLG + BA, data=moneyball)

# View model summary to check which variables are most important
summary(RunsReg)

```
**Insights**:

* **OBP (On Base Percentage)** and **SLG (Slugging %)** are *very strong predictors* of scoring runs.
* Surprisingly, **BA (Batting Average)** adds little value and even has a slightly negative effect.

```{r}
# Build a simpler regression model to predict Runs Scored
# This time removing BA since it wasn’t helpful
RunsReg = lm(RS ~ OBP + SLG, data=moneyball)

# View updated model summary
summary(RunsReg)
```
 **Conclusion**:
This simpler model still explains **93% of the variation in Runs Scored**.
This supports the Moneyball idea: **Focus on OBP and SLG** when picking players, not BA.
---

```{r}
# Build a regression model to predict Runs Allowed (RA)
# Based on the opponent's OBP and SLG (OOBP and OSLG)
RunsAllowedReg = lm(RA ~ OOBP + OSLG, data=moneyball)

# View the model summary
summary(RunsAllowedReg)
```

 **What we learn**:

* Teams allow more runs when opponents have higher OBP and SLG.
* You want pitchers and defenders who **limit how often opponents get on base or hit for power**.

---


# **In_Class _Activity8**

Exercise 1
If a baseball team’s OBP is 0.361 and  SLG is 0.409, how many runs do we expect the team to score?
Using the linear regression model constructed during the lecture (the one that uses OBP and  SLG as independent variables), find the number of runs we expect the team to score:

```{r}
# Build a regression model to predict Runs Scored
RunsReg = lm(RS ~ OBP + SLG, data=moneyball)

# View updated model summary
summary(RunsReg)
```

```{r}
runs_scored =-804.63 +2737.77*(0.361)+1584.91*(0.409)
runs_scored
```
We expect the team to score between 831 and 831 runs

---

 Exercise 2


If a baseball team’s opponents OBP (OOBP) is 0.267 and opponents SLG (OSLG) is 0.392, how many runs do we expect the team to allow?
Using the linear regression model discussed during the lecture (the one on the last slide of the previous video), find the number of runs we expect the team to allow. 

```{r}
# Build a regression model to predict Runs Allowed (RA)
# Based on the opponent's OBP and SLG (OOBP and OSLG)
RunsAllowedReg = lm(RA ~ OOBP + OSLG, data=moneyball)

# View the model summary
summary(RunsAllowedReg)
```

```{r}
runs_allowed =-837.38 +2913.60 * (0.267)+1514.29* (0.392)
runs_allowed
```
